Reproducible Research: Peer Assessment 1
=============================================

##Load and process data
```{r}
  data <- read.csv("activity.csv")
```
##Activity per day
Calculate number of steps taken per day, NA values will be ignored
```{r}
	totalsteps <- aggregate(data$steps,by=list(data$date),sum,na.rm=TRUE)
	colnames(totalsteps) <- c("Date","Total_Steps")
```
Create histogram of total steps taken per day
```{r}	
  hist(as.numeric(totalsteps$Total_Steps),main="Frequency of total steps per day",
	xlab="Total Steps",col="green")
```
Calculate and report the mean and median of total steps per day
```{r}	
  meansteps <- mean(totalsteps$Total_Steps,na.rm=TRUE)
	print(meansteps)
	mediansteps <- median(totalsteps$Total_Steps,na.rm=TRUE)
	print(mediansteps)
```
##Average Daily activity pattern
Create a time series plot of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)
```{r}
  avgsteps <- aggregate(data$steps,by=list(data$interval),mean,na.rm=TRUE)
	colnames(avgsteps) <- c("Interval","Avg_Steps")
	with(avgsteps,plot(Interval, Avg_Steps,type="l"
	    ,main="Average steps for each 5 min interval",
		  ylab="Average Steps",xlab="Interval"))
```
Calculate the 5 minute interval with the largest average steps across all days	
```{r}
  maxavgint <- subset(avgsteps,Avg_Steps == max(avgsteps$Avg_Steps))
	print(maxavgint)
```
##Imputing missing values
Calculate the number of missing values in dataset
```{r}
  totalNA <- colSums(is.na(data))
	print(totalNA)
```
Replace NA values with the mean across days for the interval with the NA rounded to a whole integer
```{r}
  imputed <- data 
	  for (i in 1:nrow(imputed)) 
	  {
    		if (is.na(imputed$steps[i])) 
		    {
        	imputed$steps[i] <- round(avgsteps[which(imputed$interval[i] 
					== avgsteps$Interval),]$Avg_Steps)
    		}
	}
	totalNAimputed <- colSums(is.na(imputed))
	print(totalNAimputed)
```
Calculate number of steps taken per day with imputed data
```{r}
  totalstepsimputed <- aggregate(imputed$steps,by=list(imputed$date),sum)
	colnames(totalstepsimputed) <- c("Date","Total_Steps")
```
Create histogram of total steps taken per day
```{r}
  hist(as.numeric(totalstepsimputed$Total_Steps),
		main="Frequency of total steps per day (imputed data)",
		xlab="Total Steps",col="blue")
```
Calculate and report the mean and median of total steps per day
```{r}
  meanstepsimputed <- mean(totalstepsimputed$Total_Steps)
	print(meanstepsimputed)
	medianstepsimputed <- median(totalstepsimputed$Total_Steps)
	print(medianstepsimputed)
```
Compare mean and median from data with missing values to data without
```{r}
  diff_mean <- abs(meansteps - meanstepsimputed)
	print(diff_mean)
	diff_median <- abs(mediansteps - medianstepsimputed)
	print(diff_median)
```
##Differences in activity between weekends and weekdays
Create new variable indicating whether the date is a weekend or weekday
```{r}
	imputed$Weekday_Weekend <- weekdays(as.Date(imputed$date,"%Y-%m-%d")) 
	imputed$Weekday_Weekend <- gsub("Monday","Weekday",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Tuesday","Weekday",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Wednesday","Weekday",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Thursday","Weekday",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Friday","Weekday",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Saturday","Weekend",imputed$Weekday_Weekend)
	imputed$Weekday_Weekend <- gsub("Sunday","Weekend",imputed$Weekday_Weekend)
```
Create a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r}
  avgsteps2 <- aggregate(imputed$steps,by=list(interval=imputed$interval,
    weekday=imputed$Weekday_Weekend),mean)
	colnames(avgsteps2) <- c("Interval","Weektime","Avg_Steps")
	library(ggplot2)
	with(avgsteps2,qplot(Interval, Avg_Steps,geom="line",facets=Weektime~.,
		main="Average steps for each interval (imputed data)",xlab="Interval",
		ylab="Average steps"))
```